Dynamic Load Balancing for\\Heterogeneous Systems

### 2011-12-05 paper/paper.tex
Improved the results section of the paper.

\section{Results and Analysis}
We tested the different benchmark applications using different hardware environments.
For the discrete GPU tests, we tested our application on an Intel Core Duo CPU with a
AMD HD5450 GPU attached via PCIe.  For the Fused CPU+GPU environment we used an AMD
Fusion A-3850.  Linux was used for the discrete environment, while Windows was used
on the Fused environment (in order to take advantage of newer drivers that allowed
for better hardware features on the new Fusion machine).

@@

First, the reduction benchmark was run on the discrete GPU platform with a variety of
different data sizes and load balancing policies, as shown in Figure \ref{fig:reduce_discrete}.
Interestingly, the CPU outperformed the GPU on this particular benchmark.  This is likely
due to the GPU not utilizing all of its stream processors during a reduction operation, as
discussed previously.  Unfortunately, none of the static load balancing schemes were able
to outperform just the CPU; all the static times were somewhere in between the CPU-only and
GPU-only times.  Additionally, the dynamic load balancing scheme performed significantly
worse than all the other schemes, and more than 100\% worse than just the CPU.
100\% worse than just the CPU.

@@

Now, consider Figure \ref{fig:vector_discrete}, where the vector add benchmark was run on
the discrete GPU platform.  In this case, we see that the GPU has nearly a 4x speedup over
just the CPU.  However, similar to the reduction benchmark, the static load balancing schemes
never outperform just the GPU alone, with all of them lying somewhere between the CPU-only
and GPU-only runs.  Also similar to the reduction benchmark, the dynamic load balancing scheme
performs significantly worse than all the other schemes.

@@

The results for the vector add+ benchmark on the discrete GPU platform are shown in Figure
\ref{fig:vector_plus_discrete}.  There are a few notable trends in this figure that are not
present in the previous two.  The first is is how drastically the GPU outperforms the CPU --
with over a 100x speedup over just the CPU with 16~MB of data.  The second is that the
dynamic load balancing drastically outperforms the static load balancing schemes.  This is
almost certainly because the GPU so drastically outperforms the CPU.  With the dynamic load
balancing scheme, only small chunks of data are given to the CPU at a time, rather than a
static percentage.

@@

For the above figures, the static load balancer was run with three different
data ratios run on the GPU -- 25\%, 50\%, and 75\%.  The general trend was that
as more of the workload was placed on the GPU, the runtime improved.  As a
result, we run the vector add benchmark again with a high granularity of data
ratios.  Figure \ref{fig:static_discrete} shows  this benchmark on the discrete
GPU platform with between 90\% and 100\% of the data on the GPU, with a 0.1\%
granularity.  The total execution time, including data transfers, has an
obvious downward trend as more of the data is executed on the GPU.  That said,
there are a couple of static ratios which yielded better performance than
executing all of the data on the GPU.  Most notably, when only 98.6\% of the
data was executed on the GPU.  In this instance, the overall execution time was
325.27 seconds, compared to 333.31 seconds with only the GPU -- a ~2.4\%
speedup.  However, these few instances of speedup may simply be due to variability.

@@

Next, consider Figures \ref{fig:reduce_fused}, \ref{fig:vector_fused}, and
\ref{fig:vector_plus_fused}. The figures demonstrate a variety of different
load balancing options for the Reduction, Vector Add, and Vector Add+
benchmarks, across data sizes ranging between 1~MB and 16~MB.  Each figure
depicts the total execution time for the CPU-only and GPU-only computations,
as well as static load balancing with 25\%, 50\%, and 75\% of the data on the
GPU.  In each of the figures, it's evident that the GPU performs significantly
faster than the CPU for the given benchmarks, with the static partitioning
schemes lying somewhere between the two.

The only deviation from the above is the reduction benchmark, in which the GPU
only performs ~14\% faster than the CPU in the 16~MB case (0.104 seconds for the
GPU, 0.091 seconds for the CPU), and some of the static load balancing schemes
perform slightly better than just the GPU.  For instance, in the 9~MB and 10~MB
cases, both the 50\% and 75\% static schemes performed better than just the GPU
(0.5\% and 4.4\% respective speedups for 9~MB, 1.5\% and 3.6\% respective speedups
for 10~MB). What is also interested to note is that on the fused CPU+GPU architecture,
the GPU outperformed the CPU on the reduction benchmark.  However, the opposite was
true of the reduction benchmark on the discrete GPU platform.

@@

The above results have lead us to believe that the CPU and GPU kernels are
not running truly in parallel as we initially believed.  As such, we used
an AMD profiler to report how long each of the kernels took using the static
load balancer with 95\% of the data on the CPU.  Figure \ref{fig:sequential_discrepancy}
shows the discrepancy between the actual total runtime of the vector add
benchmark and the expected runtime (that is, the sum of the CPU kernel runtime
and the GPU kernel runtime).  If the kernels were truly running in parallel,
we would expect to see a negative discrepancy between the actual runtime and
expected runtime.

@@

With regard to the discrete GPU platform and the fused CPU+GPU platform, it
seems that the fused platform has some advantages.  For instance, consider 
Figures \ref{fig:reduction_discrete} and \ref{fig:reduction_fused}.  Although
we will not compare overall runtimes of the two benchmarks, we can see that
the GPU of the fused platform performed better than the GPU of the discrete
platform relative to their CPUs.  That is, the CPU outperformed the GPU on
the reduction benchmark on the discrete GPU architecture, but the converse
was true of the fused platform.  This may be due to the fact that the fused
platform shares memory between the CPU and GPU, which lowers data transfer
times between the two devices.  On the discrete GPU system, the GPU typically
experienced significant data transfer times for the GPU, and the effect of
this would be minimized on the fused platform.

>>>>

\section{Results and Analysis}
We tested the different benchmark applications using different hardware environments.
For the discrete GPU tests, we tested our application on an Intel Core Duo CPU with a
AMD HD5450 GPU attached via PCIe.  For the Fused CPU+GPU environment we used an AMD
Fusion A-3850.  Linux was used for the discrete environment, while Windows was used
on the Fused environment in order to take advantage of newer drivers that allowed
for better hardware features on the new Fusion machine.

@@

First, the reduction benchmark was run on the discrete GPU platform with a variety of
different data sizes and load balancing policies, as shown in Figure \ref{fig:reduce_discrete}.
Interestingly, the CPU outperformed the GPU on this particular benchmark.  This is likely
due to the GPU not utilizing all of its stream processors during a reduction operation, as
discussed previously.  It is also the case that the large amount of synchronization that must
occur in each iteration of the GPU reduction, and the number of kernel launches required for
each iteration that is not present in the CPU reduction. In addition, because the CPU had better
performance on the kernel alone, it saw an even greater improvement when the data transfer costs
were considered, with the CPU-only case requiring no data movement at all. Unfortunately, none of
the static load balancing schemes were able to outperform the CPU-only scheme; all the static times were
somewhere in between the CPU-only and GPU-only times.  Additionally, the dynamic load balancing
scheme performed significantly worse than all the other schemes, and more than 100\%
worse than just the CPU.  The cause for this degredation in performance was likely due to the
overhead of scheduling threads and communciation between those threads, as well as moving data to
the GPU in small chunks when the GPU showed very poor performance compared to the CPU.

@@

Now, consider Figure \ref{fig:vector_discrete}, where the VectorAdd benchmark was run on
the discrete GPU platform.  In this case, we see that the GPU has nearly a 4x speedup over
just the CPU.  The data parallelism and lack of any synchronization requirements makes this
speedup over the CPU possible.  However, similar to the reduction benchmark, the static load balancing schemes
never outperform just the GPU-only scheme, with all of them lying somewhere between the CPU-only
and GPU-only runs.  Also similar to the reduction benchmark, the dynamic load balancing scheme
performs significantly worse than all the other schemes.

@@

The results for the VectorAdd+ benchmark on the discrete GPU platform are shown in Figure
\ref{fig:vector_plus_discrete}.  There are a few notable trends in this figure that are not
present in the previous two.  The first is is how drastically the GPU outperforms the CPU --
with over a 100x speedup over just the CPU with 16~MB of data.  Because of the very large
amount of computation that needs to be performed, compared to the low amount of data accesses
required in the kernel (3 data accesses for 1000 additions), the GPU is able to outperform
the CPU because of its very large throughput for computationally intensive kernels.  The second 
thing to notice for this benchmark specificially is that the
dynamic load balancing drastically outperforms the static load balancing schemes.  This is
almost certainly because the GPU so drastically outperforms the CPU.  With the dynamic load
balancing scheme, only small chunks of data are given to the CPU at a time, rather than a
static percentage, allowing for the GPU to take advantage of much more of its speedup over
the CPU. Increasing the chunk size for the GPU (taking two chunks instead of one at a time)
would likely yield in a greater speedup for the dynamic load-balancing case.

@@

For the above figures, the static load balancer was run with three different
data ratios run on the GPU -- 25\%, 50\%, and 75\%.  The general trend was that
as more of the workload was placed on the GPU, the runtime improved, with
an exception in the Reduce benchmark due to the superior CPU performance.  As a
result, we run the vector add benchmark again with a high granularity of data
ratios.  Figure \ref{fig:static_discrete} shows  this benchmark on the discrete
GPU platform with between 90\% and 100\% of the data on the GPU, with a 0.1\%
granularity.  The total execution time, including data transfers, has an
obvious downward trend as more of the data is executed on the GPU.  That said,
there are a couple of static ratios which yielded better performance than
executing all of the data on the GPU.  Most notably, when only 98.6\% of the
data was executed on the GPU.  In this instance, the overall execution time was
325.27 seconds, compared to 333.31 seconds with only the GPU -- a ~2.4\%
speedup.  However, these few instances of speedup may simply be due to variability in
the system.

@@

Next, consider Figures \ref{fig:reduce_fused}, \ref{fig:vector_fused}, and
\ref{fig:vector_plus_fused}. The figures demonstrate a variety of different
load balancing options for the Reduction, VectorAdd, and VectorAdd+
benchmarks, across data sizes ranging between 1~MB and 16~MB on a Fused CPU+GPU platform.
Each figure depicts the total execution time for the CPU-only and GPU-only computations,
as well as static load balancing with 25\%, 50\%, and 75\% of the data on the
GPU.  In each of the figures, it's evident that the GPU performs significantly
faster than the CPU for the given benchmarks, with the static partitioning
schemes lying somewhere between the two.  As you can see in the figures, while the
overall performance numbers have changed between the discrete and fused architectures
the major trends from the previous graphs still hold true.

The only deviation from the above is the reduction benchmark, in which the GPU
only performs ~14\% faster than the CPU in the 16~MB case (0.104 seconds for the
GPU, 0.091 seconds for the CPU), and some of the static load balancing schemes
perform slightly better than just the GPU.  For instance, in the 9~MB and 10~MB
cases, both the 50\% and 75\% static schemes performed better than just the GPU
(0.5\% and 4.4\% respective speedups for 9~MB, 1.5\% and 3.6\% respective speedups
for 10~MB). What is also interesting to note is that on the fused CPU+GPU architecture,
the GPU outperformed the CPU on the reduction benchmark.  However, the opposite was
true of the reduction benchmark on the discrete GPU platform.  The difference in this
performance was likely due to the increased transfer speeds between the CPU and GPU
on the fused platform.  It is also the case that the SIMD engine on the Fusion APU
is slightly more powerful than the discrete GPU that we have used for testing, allowing
kernel execution times on the APU to be lower than that of the discrete GPU.

@@

The above results have lead us to believe that the CPU and GPU kernels are
not running truly in parallel as we initially believed.  As such, we used
the OpenCL profiling capabilities to report how long each of the kernels took using the static
load balancer with 95\% of the data on the CPU.  Figure \ref{fig:sequential_discrepancy}
shows the discrepancy between the actual total runtime of the vector add
benchmark and the expected runtime of sequential calls to the kernels 
(that is, the sum of the CPU kernel runtime
and the GPU kernel runtime).  If the kernels were truly running in parallel,
we would expect to see a negative discrepancy between the actual runtime and
expected runtime.  It was impossible for us to simply time the start and end
times of the different kernels, because while that information was available
for each kernel launch, it was kept as device-speciifc times. We could not
reconcile the two different timing schemes, because of the different clock
frequencies and the lack of a mapping from time from clock start for either
device.

@@

With regard to the discrete GPU platform and the fused CPU+GPU platform, it
seems that the fused platform has some advantages.  For instance, consider 
Figures \ref{fig:reduce_discrete} and \ref{fig:reduce_fused}.  Although
we will not compare overall runtimes of the two benchmarks, we can see that
the GPU of the fused platform performed better than the GPU of the discrete
platform relative to their CPUs.  That is, the CPU outperformed the GPU on
the reduction benchmark on the discrete GPU architecture, but the converse
was true of the fused platform.  This may be due to the fact that the fused
platform shares memory between the CPU and GPU, which lowers data transfer
times between the two devices.  On the discrete GPU system, the GPU typically
experienced significant data transfer times for the GPU, and the effect of
this would be minimized on the fused platform.
